A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-14 papers/static-checker/ecoop2024/ecoop2024.tex
write the life-cycle check and other revisions

We report the foundations of the design and implementation of XXX and
conduct NNN case studies that demonstrate the expressiveness of our tool.
\end{abstract}

@@

Once a smart contract is published on a blockchain, it cannot be
changed any more. Unfortunately, there are plenty of
cases~\cite{examples-for-broken-smart-contracts} where the blockchain 
community identified vulnerabilities in smart contracts despite
extensive testing before deployment.
In addition, the novel execution model of smart contract
languages has led to unforeseen errors due to a lack of familiarity
with their intricacies. These observations along with the contract's
immutability underscore the need to ensure its correctness before deployment.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract Auction = spec 
  storage := Pair (auction_open: bool) 
                     Pair (highest_bidder: address)
                          (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance
    post-condition := post_highest_bidder = Sender 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance, highest_bidder) 
       && post_contract_owner = contract_owner ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The actual specification consists of a precondition and a
postcondition with the usual partial correctness semantics. The
postcondition refers to components of the final storage by prefixing
\lstinline|post_| to their name. The variables \lstinline|Balance|,
\lstinline|Amount|, and \lstinline|Sender| correspond to the values
returned by the like-named Michelson instructions that return the
current balance of the contract, the amount transferred with the call,
and the address of the caller (sender).

Our symbolic interpreter calculates a postcondition predicate from the specified
precondition. The SMT solver Z3 \cite{Z3} checks that this predicate
implies the specified postcondition.

@@

Listing~\ref {lst:simple-program-add} contains a short example of a Michelson program.
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code { 
(*               pair int int :: [] *)
   DUP; UNPAIR;
(* int :: int :: pair int int :: [] *)
   COMPARE; GE; 
(*       bool :: pair int int :: [] *)
   IF {UNPAIR; ADD} {PUSH string 'Unexpected Pair'; FAILWITH};
(*                        int :: [] *)
   NIL operation; PAIR
}
\end{lstlisting}
This program first duplicates its input (\lstinline|DUP|), decomposes
the pair $P\times S$ into its components (\lstinline|UNPAIR|),
compares the two value on top of the stack (\lstinline|COMPARE|), and
transforms the result into a boolean (\lstinline|GE|). The
\lstinline|IF| instruction pops the boolean and either executes the
first (for true) or second (for false) list of instructions. By now
only the initial pair is on the stack. In the true branch,
\lstinline|UNPAIR| decomposes, and then \lstinline|ADD| adds the
components. In the false branch, the \lstinline|PUSH| instruction pushes
a string message on the stack. The \lstinline|FAILWITH| instruction
abort the contract execution: tokens are consumed, but the storage and
the balance of the contract remain unchanged.

@@

One of the main challenges with formal verification is that users are
often required to express  system properties in a specialized
specification language. This requirement poses two problems: it
demands expertise in formal verification from the specifier and
introduces a gap between the actual code and its specification. In our
framework, users are not obliged to possess formal verification
expertise. A basic understanding of Michelson data types and their
usage when interacting with a smart contract on the Tezos blockchain
is sufficient.

\begin{figure}[tp]
\begin{align*}
\textbf{c} & ::= 
    \text{Balance}
   \Mid \text{post\_Balance}
   \Mid \text{Amount}
   \Mid \text{Sender}
   \Mid \text{Source} 
   \Mid \text{Now}
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id}
   \Mid \text{Self\_address}
   \Mid \text{Total\_voting\_power}
   \Mid \text{Voting\_power} \\
\textbf{e} & :: =  \textbf{i} \Mid  \textbf{s}  \Mid  \textbf{b} \Mid \textbf{c} \Mid (\text{x} : \text{T}) \\
&\Mid \text{Unit}  \Mid \text{Never}  \Mid \text{True}  \Mid \text{False} \\
&\Mid \text{Pair } \textbf{e } \textbf{e}  \Mid \text{Left } \textbf{e}  \Mid \text{Right } \textbf{e}  \Mid \text{Some } \textbf{e} \Mid \text{None T}  \\
& \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ \/ } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e}, \textbf{e}, \textbf{e}) 
 \Mid \text{get\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{get\_big\_map} (\textbf{e}, \textbf{e}) \\
& \Mid \text{get\_contract} (\textbf{e}, \textbf{e})  
 \Mid \text{abs} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{$\wedge$} \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e}
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
&\Mid \text{forall} (\textbf{e}, \textbf{p})
                                            \Mid \text{exist} (\textbf{e}, \textbf{p})
                                            \Mid \textbf{p} \text{ and } \textbf{p}\Mid \textbf{p} \text{ or } \textbf{p}  \Mid \textbf{p} \text{ nor } \textbf{p} \Mid  \text{ not } \textbf{p} \Mid \textbf{p} \text{ imply } \textbf{p} \\
\end{align*}
\caption{Syntax of predicates, p}
\label{fig:syntax-predicates}
\end{figure}

\begin{lstlisting}[float=tp,captionpos=b,caption={Smart contract property specification},label={lst:smart-contract-property-specification},numbers=left]
mcontract Name = spec
  code := {...}
  input := t
  output := t'
  pre-condition := cd
  post-condition := cd'
\end{lstlisting}

@@

Listing~\ref{lst:smart-contract-property-specification} shows the
syntax of our property specification language. Here, \KMCONTRACT,
\KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are
keywords. The \KNAME\ is replaced by the contract name. In the input
and output fields, the input and output patterns are specified.
\atha{What patterns are supported? Patterns are not explained!}
Since
Michelson operates on a stack, we assume that the stack has only one
element, as specified in the input pattern. Similarly, the output
stack is assumed to be a stack with only one element, which is
specified in the output field. Thus, the syntax allows us to specify
incomplete contracts, too.
\pt{I disagree (though it would be a useful feature). To this end, there should be patterns to cover stacks
  with more than one element. As it stands, the input pattern can only
speak about the top of the stack and the output pattern only about the
second component of the top of the stack, which is assumed to be a
pair.\\
Moreover, a specification might want to say something about the list
of operations, too, e.g., that tokens are transferred to some address.}
For a complete smart contract, the output
should be a pair consisting of an operation list and a storage.

For verification purposes we elide the operation list and represent
the output solely by the pattern of the value in the storage after the
execution concludes. In some cases, the output pattern can be omitted
for a complete contract, given that the input pattern must be a pair
of parameters and storage. Hence the storage pattern is part of the
input pattern. In such cases, the prefix \lstinline/post_/ is used
before the variables from the input term to indicate the corresponding
storage value after the execution.

The input and output patterns represent Michelson terms, which can be
either concrete values or symbolic values, including variables. When a
variable appears for the first time, it needs to be accompanied by its
corresponding type declaration. Subsequent mentions can omit the type
declaration.  Formally, the pre- and postcondition instances are
formulas in first-order logic that can be handled by an SMT solver (in
our case, Z3). The specification language is constituted of predicates
defined using the syntax defined in Figure~\ref{fig:syntax-predicates}. Formulas that are used in assertions or
added to solvers are terms of sort Boolean. Otherwise, terms of
Boolean and non-Boolean sort may be mixed in any combination where
sorts match up. Universal and existential quantifiers bind variables
to the scope of the quantified formula.
\atha{More explanation is needed about expressions and formulas. Why
  do you support left and right, but not lists? Why Never? What is the
meaning of transfer\_token, get\_map, get\_contract? The constants
must be explained (though you could do that in
Section~\ref{sec:background}). (Less common items could be omitted!)}

\begin{lstlisting}[float,captionpos=b,caption={Specification of Add contract},label={lst:add-contract-specification},numbers=left]
mcontract Add = spec
  code := {UNPAIR; COMPARE; GE; 
          IF {ADD}{PUSH string 'Unexpected Pair'; FAILWITH}}
  input := Pair (x : int) (y : int)
  output := (z : int)
  pre-condition := (x >= 5) | (y > 3)
  post-condition := (z >= 8)
\end{lstlisting}

Let us consider the contract \texttt{Add}  in Listing
\ref{lst:add-contract-specification} as an   example of a property
specification.  Here, the program takes as input a stack with only one
element, which is a pair of two integers, symbolically specified in
the \lstinline|input| field. The program unpairs the input pair to
obtain two elements as two integers \lstinline/x/ and \lstinline/y/.
Next, the program performs a comparison between
these two integers. If \lstinline/x/ is greater than or equal to
\lstinline/y/, the program constructs a new stack with a solitary
element at the top representing the sum of  \lstinline/x/ and
\lstinline/y/, designated as \lstinline/z/ by the \lstinline|output| directive. Otherwise, the execution
terminates with an error message containing the content
\lstinline|'Unexpected pair'|.
\atha{This Michelson code with its explanation might be an example for Section~\ref{sec:background}}
The specification continues by defining
pre- and postconditions, where the precondition states the assumption
that the integer \lstinline/x/ is greater than or equal to 5 and \lstinline/y/
is greater than or equal to 3. The symbol  \lstinline/|/ stands for
logical disjunction. The postcondition asserts that the result
\lstinline/z/ is greater than or equal to 8.
\atha{This specification is meaningless. Why not use $x\ge y$ as
  precondition and $z = x+y$ as postcondition?}

@@

In XXX, we can specify several entrypoints at the same
time. This design enables us to specify and verify all entrypoints
at once and also to describe and verify relationships between
entrypoints.  The syntax is shown in Listing~\ref {lst:multiple-entrypoint-specification},
\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax},label={lst:multiple-entrypoint-specification},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      code := {c2}
      input := t2
      output := t2'
      pre-condition := cd2
      post-condition := cd2'
   entrypoint %c 
      ...
   <entrypoint relations>
\end{lstlisting}
where the entrypoint name is written after the \lstinline/%/ symbol
and is followed by a property specification for the entrypoint. The
specification finishes with specifying entrypoint relations,
which we explain next.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := Pair (auction_open: bool) 
                     Pair (highest_bidder: address)
                          (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance
    post-condition := post_highest_bidder = Sender 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance, highest_bidder) 
       && post_contract_owner = contract_owner ;

  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open 
       && Sender = contract_owner                
    post-condition := not post_auction_open 
       && post_Balance = 0 
       && transfer_token (Unit, Balance, contract_owner) ;

    (%create -> %bid) with (auction_open = true)  
       && (Amount > Balance) 
    | (%create  -> %close) with (auction_open = true) 
       && (Sender = contract_owner)
    | (%bid -> %bid) with (Amount > Balance)  
    | (%bid -> %close) with (Sender = contract_owner)  
    | not (%close -> %bid) 
    | not (%close -> %close)
\end{lstlisting}

There are several properties that we want to guarantee for this
contract. Let us have a look at the specification for the auction
contract in Listing \ref{lst:auction-contract-specification}. We omit
the Michelson code of the smart contract due to its length.  For the
\lstinline/%bid/ entrypoint, the input stack has only one element,
which is a pair consisting of the parameter and the current
storage. The parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is represented as a nested pair with the
structure described by the \lstinline|storage| phrase. 
The output term should maintain the required pattern of the storage,
but the values may be changed and these changes are recorded in
different symbolic variables. In this specification, we add the prefix
\lstinline/post_/ to the variable names to refer to the storage
returned by the program. The output pattern could be safely removed in
this example because the tool automatically generates an output
pattern corresponding the one specified here.

@@

We verify the contract against its specifications in the contract
module by symbolic execution. Symbolic execution involves calculating
a symbolic stack at each transition from one Michelson instruction to
the next. The symbolic interpreter is fully typed and rejects
ill-typed Michelson programs, provided the ill-typed part is reachable
from the default entry point.  The initial stack is derived from the
input pattern for each the entry point. To maintain symbolic values as
concrete as possible, some instructions may be deemed unreachable or
only reachable under specific conditions. This process yields, for
each entry point, a set of final symbolic stacks accompanied by a path
condition indicating when this final state is reachable. A state is
discarded from the final state space if the path condition is not
satisfiable, indicating that the path is unreachable. Additionally,
for each \FAILWITH\ instruction, we obtain a symbolic value for the
reported message and a path condition indicating the reachability of
this instruction. We employ the SMT solver Z3  to assess the
satisfiability of a path condition and to check smart contract
properies specified by pre- and post conditions.

The architecture of the static checker is illustrated in
Figure~\ref{fig:architecture-of-static-checker}. The system comprises
three phases. The first phase is user specification, where users
provide the Michelson code of the contract and define the properties
that the smart contract should satisfy. Subsequently, the symbolic
interpreter is run on the code with the provided input pattern.  The
path conditions are then fed into a Z3 converter, which transforms
symbolic terms into Z3 formulas. Concurrently, the property
specification, specifically pre- and post conditions, also undergoes
conversion to Z3 formulas. The results of the symbolic execution and
Z3 formulas are then input into the static checker, which verifies
whether these properties hold for the smart contract
implementation. This constitutes the verification phase.

@@

In the user specification phase, users compose the specification in
our domain-specific language using the syntax outlined in
Section~\ref{sec:domain-specific-language}. Michelson code, whether
composed by users, obtained from the Tezos blockchain, or compiled
from higher-level languages, can be seamlessly used within our
framework. The specification is subsequently parsed into an Abstract
Syntax Tree (AST) in OCaml. The code and the input pattern are then
input to the symbolic interpreter. In the next phase, the pre- and
post-predicates are transformed into Z3 formulas by the Z3 converter.

>>>>

We report the foundations of the design and implementation of XXX and
conduct two case studies that demonstrate the expressiveness of our tool.
\end{abstract}

@@

Once a smart contract is published on a blockchain, it cannot be
changed any more. Unfortunately, there are plenty of
cases~\cite{dao,wallethack} where the blockchain 
community identified vulnerabilities in smart contracts despite
extensive testing before deployment.
In addition, the novel execution model of smart contract
languages has led to unforeseen errors due to a lack of familiarity
with their intricacies. These observations along with the contract's
immutability underscore the need to ensure its correctness before deployment.

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                  Pair (highest_bidder: address)
                       (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance - Amount
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance - Amount, c) ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The actual specification consists of a precondition and a
postcondition with the usual partial correctness semantics. The
postcondition refers to components of the final storage by prefixing
\lstinline|post_| to their name. The variables \lstinline|Balance|,
\lstinline|Amount|, and \lstinline|Sender| correspond to the values
returned by the like-named Michelson instructions that return the
current balance of the contract, the amount transferred with the call,
and the address of the caller (sender).

Our symbolic interpreter calculates a postcondition predicate from the specified
precondition. The SMT solver Z3 \cite{z3} checks that this predicate
implies the specified postcondition.

@@

Listing~\ref {lst:simple-program-add} contains a short example of a Michelson program.
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code { 
(*               pair int int :: [] *)
   DUP; UNPAIR;
(* int :: int :: pair int int :: [] *)
   COMPARE; GE; 
(*       bool :: pair int int :: [] *)
   IF {UNPAIR; SUB} {PUSH string 'Unexpected Pair'; FAILWITH};
(*                        int :: [] *)
   NIL operation; PAIR
}
\end{lstlisting}
This program first duplicates its input (\lstinline|DUP|), decomposes
the pair $P\times S$ into its components (\lstinline|UNPAIR|),
compares the two value on top of the stack (\lstinline|COMPARE|), and
transforms the result into a boolean (\lstinline|GE|). The
\lstinline|IF| instruction pops the boolean and either executes the
first (for true) or second (for false) list of instructions. By now
only the initial pair is on the stack. In the true branch,
\lstinline|UNPAIR| decomposes, and then \lstinline|SUB| subtract the two 
components. In the false branch, the \lstinline|PUSH| instruction pushes
a string message on the stack. The \lstinline|FAILWITH| instruction
abort the contract execution: tokens are consumed, but the storage and
the balance of the contract remain unchanged.

@@

One of the main challenges with formal verification is that users are
often required to express  system properties in a specialized
specification language. This requirement poses two problems: it
demands expertise in formal verification from the specifier and
introduces a gap between the actual code and its specification. In our
framework, users are not obliged to possess formal verification
expertise. A basic understanding of Michelson data types and their
usage when interacting with a smart contract on the Tezos blockchain
is sufficient.
\begin{figure}[tp]
\begin{align*}
\textbf{cpt} &::= 
   \text{unit} 
   \Mid\ \text{never} 
   \Mid\ \text{bool} 
   \Mid\ \text{int}
   \Mid\ \text{nat}
   \Mid\ \text{string}
   \Mid\ \text{chain-id}
   \Mid\ \text{bytes}
   \Mid\ \text{mutez} \\
   &\Mid\ \text{key-hash}
   \Mid\ \text{key}
   \Mid\ \text{signature}
   \Mid\ \text{timestamp}
   \Mid\ \text{address} \\
   &\Mid\ \text{option}\ \textbf{cpt}
   \Mid\ \text{or}\ \textbf{cpt}\ \textbf{cpt}
   \Mid\ \text{pair}\ \textbf{cpt}\ \textbf{cpt} \\
\text{T, U} &::= 
    \text{cpt}
   \Mid\ \text{option}\ \text{T}
   \Mid\ \text{list}\ \text{T}
   \Mid\ \text{set}\ \textbf{cpt} 
   \Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \text{T}
   \Mid\ \text{ticket}\ \textbf{cpt} 
   \Mid\ \text{pair}\ \text{T}\ \text{U}
   \Mid\ \text{or}\ \text{T}\ \text{U}
   \Mid\ \text{lambda}\ \text{T}\ \text{U} \\
   &\Mid\ \text{map}\ \textbf{cpt}\ \text{T}
   \Mid\ \text{big-map}\ \textbf{cpt}\ \text{T}
   \Mid\ \text{chest} 
   \Mid\ \text{chest-key}
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}

\begin{figure}[tp]
\begin{align*}
\textbf{c} & ::= 
    \text{Balance}
   \Mid \text{post\_Balance}
   \Mid \text{Amount}
   \Mid \text{Sender}
   \Mid \text{Source} 
   \Mid \text{Now}
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id}
   \Mid \text{Self\_address}
   \Mid \text{Total\_voting\_power}
   \Mid \text{Voting\_power} \\
\textbf{t} &::= 
   \textbf{n}: \text{nat}
   \Mid\ \textbf{i}: \text{int}
   \Mid\  \textbf{s}: \text{string} 
   \Mid\  \textbf{b}: \text{bytes}
   \Mid\  \textbf{c}  
   \Mid\  \text{x} 
   \Mid\ \text{x}: \text{T} \\
   &\Mid\ \UNIT 
   \Mid\ \NEVER 
   \Mid\ \TRUE 
   \Mid\ \FALSE 
   \Mid\ \PAIR\ \text{t1 t2} \\
   &\Mid\ \LEFT\ \text{t}\  \text{T}
   \Mid\ \RIGHT\ \text{T}\ \text{t}
   \Mid\ \SOME\ \text{t}
   \Mid\ \NONE\ \text{T} 
   \Mid\ \text{\{t ; ... \}}: \text{list T}
   \Mid\ \text{\{t ; ... \}}: \text{set T} \\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{map T U}
   \Mid\ \text{\{ Elt t1 t2 ; ... \}}: \text{big\_map T U} \\
\textbf{e} & :: = \textbf{t}  \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ / } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e}, \textbf{e}, \textbf{e}) 
 \Mid \text{get\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{get\_big\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{check\_mem} (\textbf{e}, \textbf{e})  \\
& \Mid \text{get\_contract} (\textbf{e}, \textbf{e})  
 \Mid \text{contain} (\textbf{e}, \textbf{e})  
 \Mid \text{get\_substring} (\textbf{e}, \textbf{e}, \textbf{e})  
 \Mid \text{check\_sig} (\textbf{e}, \textbf{e}, \textbf{e})  \\
& \Mid \text{abs} (\textbf{e})  
 \Mid \text{is\_nat} (\textbf{e})  
 \Mid \text{to\_int} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{cons} (\textbf{e}, \textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{ $\wedge$ } \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e}
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
&\Mid \text{forall} (\textbf{e}, \textbf{p})
\Mid \text{exist} (\textbf{e}, \textbf{p})
\Mid \textbf{p} \text{ and } \textbf{p}\Mid \textbf{p} \text{ or } \textbf{p}  \Mid \textbf{p} \text{ nor } \textbf{p} \Mid  \text{ not } \textbf{p} \Mid \textbf{p} \text{ imply } \textbf{p} \\
\end{align*}
\caption{Syntax of terms, t and predicates, p}
\label{fig:syntax-terms-predicates}
\end{figure}

\begin{lstlisting}[float=tp,captionpos=b,caption={Smart contract property specification for a simple contract},label={lst:smart-contract-property-specification},numbers=left]
mcontract name = spec
  code := {...}
  input := t
  output := t'
  pre-condition := cd
  post-condition := cd'
\end{lstlisting}

@@

Listing~\ref{lst:smart-contract-property-specification} shows the
syntax of our property specification language for a simple contract. Here, \KMCONTRACT,
\KSPEC, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are
keywords. The \KNAME\ is replaced by the contract name. In the input
and output fields, the input and output terms are specified. 
\atha{What terms are supported? terms are not explained!}
Since
Michelson operates on a stack, we assume that the stack has only one
element, as specified in the input term. Similarly, the output
stack is assumed to be a stack with only one element, which is
specified in the output field. Thus, the syntax allows us to specify
incomplete contracts, too.
\pt{I disagree (though it would be a useful feature). To this end, there should be terms to cover stacks
  with more than one element. As it stands, the input term can only
speak about the top of the stack and the output term only about the
second component of the top of the stack, which is assumed to be a
pair.\\
Moreover, a specification might want to say something about the list
of operations, too, e.g., that tokens are transferred to some address.}

For a complete smart contract, the output should be a pair consisting of an operation list and the storage's state after the completion of the code.  For verification purposes we elide the operation list and represent the output solely by the term of the value in the storage after the execution concludes. In some cases, the output term can be omitted
for a complete contract, given that the input term must be a pair
of parameters and storage. Hence the storage term is part of the
input term. In such cases, the prefix \lstinline/post_/ is used
before the variables from the input term to indicate the corresponding
storage value after the execution.

The input and output terms represent Michelson terms, which can be
either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its
corresponding type declaration. Subsequent mentions can omit the type
declaration.  Formally, the pre- and postcondition instances are
formulas in first-order logic that can be handled by an SMT solver (in
our case, Z3). The specification language is constituted of predicates.  The syntax terms and  predicates are defined in Figure~\ref{fig:syntax-terms-predicates}. Formulas that are used in assertions or
added to solvers are terms of sort Boolean. Otherwise, terms of
Boolean and non-Boolean sort may be mixed in any combination where
sorts match up. Universal and existential quantifiers bind variables
to the scope of the quantified formula.
\atha{More explanation is needed about expressions and formulas. Why
  do you support left and right, but not lists? Why Never? What is the
meaning of transfer\_token, get\_map, get\_contract? The constants
must be explained (though you could do that in
Section~\ref{sec:background}). (Less common items could be omitted!)}

\begin{lstlisting}[float,captionpos=b,caption={Specification of Sub contract},label={lst:sub-contract-specification},numbers=left]
mcontract sub = spec
  code := {DUP; UNPAIR; COMPARE; GE; IF {UNPAIR; SUB} 
          {PUSH string 'Unexpected Pair'; FAILWITH}}
  input := Pair (x : int) (y : int)
  output := (z : int)
  pre-condition := (x >= y)
  post-condition := (z = x - y) && (z >= 0)
\end{lstlisting}

Let us consider the contract \texttt{Sub}  in Listing
\ref{lst:sub-contract-specification} as an example of a property
specification.  Here, the program takes as input a stack with only one
element, which is a pair of two integers, symbolically specified in
the \lstinline|input| field. The program first duplicates the pair. It then unpairs the input pair to obtain two elements as two integers \lstinline/x/ and \lstinline/y/.
Next, the program performs a comparison between
these two integers. If \lstinline/x/ is greater than or equal to
\lstinline/y/, the program constructs a new stack with a solitary
element at the top representing the subtract of  \lstinline/x/ and
\lstinline/y/, designated as \lstinline/z/ by the \lstinline|output| directive. Otherwise, the execution
terminates with an error message containing the content
\lstinline|'Unexpected pair'|.
\atha{This Michelson code with its explanation might be an example for Section~\ref{sec:background}}
The specification continues by defining
pre- and postconditions, where the precondition states the assumption
that the integer \lstinline/x/ is greater than or equal to  \lstinline/y/. The postcondition asserts that the result
\lstinline/z/ is greater than or equal to the substract of \lstinline/x/ and \lstinline/y/, which is greater or equal to zero. The symbol \lstinline/&&/ stands for logical conjunction.
\atha{This specification is meaningless. Why not use $x\ge y$ as
  precondition and $z = x+y$ as postcondition?}

@@

In XXX, we can specify several entrypoints at the same
time. This design enables us to specify and verify all entrypoints
at once and also to describe and verify relationships between
entrypoints.  There are two options of the syntax as shown in Listing~\ref {lst:multiple-entrypoint-specification-1} and Listing~\ref {lst:multiple-entrypoint-specification-2},
where the entrypoint name is written after the \lstinline/%/ symbol
and is followed by a property specification for the entrypoint. The
specification finishes with specifying entrypoint relations,
which we explain next.

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 1)},label={lst:multiple-entrypoint-specification-1},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax (option 2)},label={lst:multiple-entrypoint-specification-2},numbers=left][h]
mcontract Name = spec
   storage := t1
   entrypoint %a 
      code := {c1}
      parameter := t2
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      ...
   <entrypoint relations>
\end{lstlisting}

@@

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract auction = spec 
  storage := Pair (auction_open: bool) 
                   Pair (highest_bidder: address)
                        (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance - Amount
       && get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
    post-condition := auction_open
       && post_highest_bidder = Sender 
       && post_contract_owner = contract_owner 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance - Amount, c);

  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open 
       && Sender = contract_owner  
       && get_contract (unit, contract_owner) = 
          Some (c: contract unit)              
    post-condition := not post_auction_open 
       && post_Balance = 0 
       && transfer_token (Unit, Balance, c)

  (%create -> %bid) with (auction_open = true)  
       and (Amount > Balance) 
       and get_contract (unit, highest_bidder) = 
           Some (c: contract unit)
  | (%create  -> %close) with (auction_open = true) 
       and (Sender = contract_owner)
       and get_contract (unit, contract_owner) = 
          Some (c: contract unit)
  | (%bid -> %bid) with (Amount > Balance - Amount) 
       and get_contract (unit, highest_bidder) = 
          Some (c: contract unit)
  | (%bid -> %close) with (Sender = contract_owner) 
       and get_contract (unit, contract_owner) = 
            Some (c: contract unit)
  | not (%close -> %bid)
  | not (%close -> %close)
\end{lstlisting}

There are several properties that we want to guarantee for this
contract. Let us have a look at the specification for the auction
contract in Listing \ref{lst:auction-contract-specification}. We omit
the Michelson code of the smart contract due to its length.  For the
\lstinline/%bid/ entrypoint, the input stack has only one element,
which is a pair consisting of the parameter and the current
storage. The parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is represented as a nested pair with the
structure described by the \lstinline|storage| phrase. 
The output term should maintain the required term of the storage,
but the values may be changed and these changes are recorded in
different symbolic variables. In this specification, we add the prefix
\lstinline/post_/ to the variable names to refer to the storage
returned by the program. The output term could be safely removed in
this example because the tool automatically generates an output
term corresponding the one specified here.

@@

We verify the contract against its specifications in the contract
module by symbolic execution. Symbolic execution involves calculating
a symbolic stack at each transition from one Michelson instruction to
the next. The symbolic interpreter is fully typed and rejects
ill-typed Michelson programs, provided the ill-typed part is reachable
from the default entry point.  The initial stack is derived from the
input term for each the entry point. To maintain symbolic values as
concrete as possible, some instructions may be deemed unreachable or
only reachable under specific conditions. This process yields, for
each entry point, a set of final symbolic stacks accompanied by a path
condition indicating when this final state is reachable. A state is
discarded from the final state space if the path condition is not
satisfiable, indicating that the path is unreachable. Additionally,
for each \FAILWITH\ instruction, we obtain a symbolic value for the
reported message and a path condition indicating the reachability of
this instruction. We employ the SMT solver Z3  to assess the
satisfiability of a path condition and to check smart contract
properies specified by pre- and post conditions.

The architecture of the static checker is illustrated in
Figure~\ref{fig:architecture-of-static-checker}. The system comprises
three phases. The first phase is user specification, where users
provide the Michelson code of the contract and define the properties
that the smart contract should satisfy. Subsequently, the symbolic
interpreter is run on the code with the provided input term.  The
path conditions are then fed into a Z3 converter, which transforms
symbolic terms into Z3 formulas. Concurrently, the property
specification, specifically pre- and post conditions, also undergoes
conversion to Z3 formulas. The results of the symbolic execution and
Z3 formulas are then input into the static checker, which verifies
whether these properties hold for the smart contract
implementation. This constitutes the verification phase.

@@

In the user specification phase, users compose the specification in
our domain-specific language using the syntax outlined in
Section~\ref{sec:domain-specific-language}. Michelson code, whether
composed by users, obtained from the Tezos blockchain, or compiled
from higher-level languages, can be seamlessly used within our
framework. The specification is subsequently parsed into an Abstract
Syntax Tree (AST) in OCaml. The code and the input term are then
input to the symbolic interpreter. In the next phase, the pre- and
post-predicates are transformed into Z3 formulas by the Z3 converter.
